Cover Times of Random Walks

### 2019-04-20 paper.tex
change wording
distribution

\DeclareMathOperator{\E}{\mathrm{E}}		     % expected value
\DeclareMathOperator{\pr}{\mathrm{P}}		     % expected value
\DeclareMathOperator{\cov}{\uptau_\textrm{cov}}  % cover time variable
\DeclareMathOperator{\tcov}{t_\textrm{cov}}      % cover time
\DeclareMathOperator{\hit}{t_{\textrm{hit}}}     % hitting time
\DeclareMathOperator{\tmina}{t_\textrm{min}^A}   % minimum time on A
\DeclareMathOperator{\Ceff}{C_\textrm{eff}}      % effective conductance
\DeclareMathOperator{\Reff}{R_{\textrm{eff}}}    % effective resistance
\DeclareMathOperator{\Rxy}{R_{\textrm{x,y}}}     % effective resistance between x,y
\DeclareMathOperator{\T}{\mathcal{T}}            % spanning tree

@@

Let $T_{j}$ be the number of steps until the first visit to $j$.
Formally,
\begin{align}
T_{j} &= \textrm{min} \left\{t > 0 : X_t = j \right\}. \nonumber
\end{align}
Let the hitting time $\hit$ be the maximum expected time between two vertices on $G$.
Note that $\E_i[T_j] = \E[T_j | X_0 = i]$.
Formally, 
\begin{align}
\hit &= \max_{i,j \in V(G)} \E_i[T_j]. \nonumber
\end{align}
Let the cover time variable $\cov$ be the first time all vertices have been
visited by the random walk $(X_t)$.
Formally, $\cov$ is the minimum time that for all $i \in V(G)$
there exists $t \leq \cov$ such that $X_t = i$.
We also define the cover time as the expected value of $\cov$ from
the worst initial vertex.
Formally,
\begin{align}
\tcov &= \max_{j \in V(G)} \E_j[\cov] . \nonumber
\end{align}

@@

\section{Distributional Aspects}
\subsection{Complete Graph}
\cite{Du11}
Theorem 2.2.4 -> Example 2.2.5 -> Theorem 3.6.5 -> Example 3.6.6

\todo{Theorem 2.2.4}

\todo{Example 2.2.5}

\todo{Theorem 3.6.5}

\todo{Example 3.6.6}

>>>>

\DeclareMathOperator{\E}{\mathrm{E}}		     % expected value
\DeclareMathOperator{\Var}{\mathrm{Var}}         % variance
\DeclareMathOperator{\I}{\mathbb{1}}         % variance
\DeclareMathOperator{\pr}{\mathrm{P}}		     % expected value
\DeclareMathOperator{\cov}{\uptau_\textrm{cov}}  % cover time variable
\DeclareMathOperator{\tcov}{t_\textrm{cov}}      % cover time
\DeclareMathOperator{\hit}{t_{\textrm{hit}}}     % hitting time
\DeclareMathOperator{\tmina}{t_\textrm{min}^A}   % minimum time on A
\DeclareMathOperator{\Ceff}{C_\textrm{eff}}      % effective conductance
\DeclareMathOperator{\Reff}{R_{\textrm{eff}}}    % effective resistance
\DeclareMathOperator{\Rxy}{R_{\textrm{x,y}}}     % effective resistance between x,y
\DeclareMathOperator{\T}{\mathcal{T}}            % spanning tree

@@

Let $T_{j}$ be the number of steps until the first visit to $j$.
Formally,
\begin{align}
T_{j} &= \textrm{min} \left\{t > 0 : X_t = j \right\}. \nonumber
\end{align}
Let the hitting time $\hit$ be the maximum expected time between two vertices on $G$.
Note that $\E_i[T_j] = \E[T_j | X_0 = i]$.
Formally, 
\begin{align}
\hit &= \max_{i,j \in V(G)} \E_i[T_j]. \nonumber
\end{align}
Let the cover time variable $\cov$ be the first time all vertices have been
visited by the random walk $(X_t)$.
Formally, $\cov$ is the minimum time that for all $i \in V(G)$
there exists $t \leq \cov$ such that $X_t = i$.
We define the cover time as the expected value of $\cov$ from
the worst initial vertex.
Formally,
\begin{align}
\tcov &= \max_{j \in V(G)} \E_j[\cov] . \nonumber
\end{align}

@@

\section{Distributional Aspects}
\cite{Du11}
Consider a random walk on the complete graph with self-edges.
Call $r$ the number of steps since we began the random walk.
Let $A_i$ be the event that vertex $i$ is not visited and
$N_n$ the number of unvisited vertices.
Then
\begin{align}
P(A_i) = (1-1/n)^r \nonumber
\end{align}
and
\begin{align}
\E[N_n] = n(1-1/n)^r. \nonumber
\end{align}

If $r/n \rightarrow c$, then
\begin{align}
\E[N_n]/n &= (1-1/n)^r = [(1-1/n)^n]^{r/n} \rightarrow (e^{-1})^c \nonumber \\
\E[N_n]/n &\rightarrow e^{-c}. \nonumber
\end{align}

We compute the variance of $N_n$.
Let $\I_m$ be the indicator random variable for $A_m$.
First find $E[N_n^2]$,
\begin{align}
\E[N_n^2] &= \E \left[ \left(\sum_{m=1}^n \I_m \right) ^2 \right] =
\E \left[ \sum_{m=1}^n \sum_{k=1}^n \I_m \I_k \right] =
\sum_{1 \leq k, m \leq n} \pr(A_k \cap A_m). \nonumber
\end{align}

Then
\begin{align}
\Var[N_n] &= \E[N_n^2] - \E[N_n]^2 \nonumber \\
&= \sum_{1 \leq k, m \leq n} \pr(A_k \cap A_m) - P(A_k)P(A_m) \nonumber \\
&= n(n-1) \{ \pr(A_k \cap A_m) - P(A_k)P(A_m) \} + 
n \{ \pr(A_k \cap A_m) - P(A_k)P(A_m) \} \nonumber \\
&= n(n-1) \{ (1-2/n)^r - (1-1/n)^{2r} \} + 
n \{ (1-1/n)^r  - (1-1/n)^{2r} \}. \nonumber
\end{align}
where $k\neq m$ in the first term and $k=m$ in the second.

As $n \rightarrow \infty$,
\begin{align}
\Var(N_n/n) &= \Var(N_n) / n^2 \nonumber\\
&\rightarrow (1-2/n)^r - (1-1/n)^{2r} = e^{-2c} - (e^{-c})^2 = 0. \nonumber
\end{align}

By Chebyshev's Inequality,
\begin{align}
\pr \left( |N_n/n - \E[N_n/n]| \geq \epsilon \right) &\leq \Var(N_n/n) / \epsilon^2
\rightarrow 0. \nonumber
\end{align}

Thus we conclude $N_n/n \rightarrow \E[N_n/n] = e^{-c}$ in probability.

The Poisson approximation of the binomial implies that if $n \rightarrow \infty$
and $r/n \rightarrow c$, then the number of visits to each vertex will approach
a Poisson distribution with mean $c$.
Then it intuitively follows that the proportion of visits to each vertex
approaches $e^{-c}$.

\begin{theorem}\label{thm:pois}
If $ne^{-r/n} \rightarrow \lambda \in [0, \infty)$, then the number of unvisited
vertices approaches a Poisson distribution with mean $\lambda$.
\end{theorem}

\begin{proof}
Observe that the probability that there are $k$ unvisited vertices
\begin{align}
(1-k/n)^r \nonumber.
\end{align}

Define $p_m(r,n)$ as the probability that exactly $m$ vertices are unvisited
when $r$ steps have been taken.
Then the probability that all vertices have been visited is the complement
of the probability that at least one vertex has not been visited.

By the principle of inclusion-exclusion,
\begin{align}
p_0(r,n) &= {n \choose 0} 1^r - {n \choose 1} (1-1/n)^r + ... \nonumber\\
&= \sum_{k=0}^n (-1)^k {n \choose k} (1-k/n)^r. \nonumber
\end{align}

If we consider the locations of the unvisited vertices,
\begin{align}
p_m(r,n) = {n \choose m} \left(1-\frac{m}{n}\right)^r p_0(r, n-m). \nonumber
\end{align}

We want to show that 
\begin{align}
{n \choose m} \left(1-\frac{m}{n}\right)^r \rightarrow \lambda ^m /m! \nonumber
\end{align}

We begin with the more straightforward upper bound.
Notice that $(1-x) \leq e^{-x}$ and recall our assumption that
$ne^{r/n} \rightarrow \lambda$.
Then
\begin{align}
{n \choose m} \left(1-\frac{m}{n}\right)^r &\leq \frac{n^m}{m!} e^{mr/n} \nonumber \\
&= \frac{(n e^{r/n})^m} {m!} \nonumber \\
&\rightarrow \lambda ^m / m! \nonumber
\end{align}

We now turn to the more tricky lower bound.
Notice that ${n \choose m} \geq (n-m)^m / m!$
\todo{why?}
Then
\begin{align}
{n \choose m} \left(1-\frac{m}{n}\right)^r &\geq \left(1-\frac{m}{n}\right)^r (n-m)^m / m! \nonumber\\
&\geq n^m (1 - \frac{m}{n})^r (1- \frac{m}{n})^m  / m! \nonumber
\end{align}

As $n \rightarrow \infty$, $\left(1-\frac{m}{n}\right)^m \rightarrow 1$ and $1/m!$ is constant.
We use the Taylor series of $\log(1-t)$ for $0 \leq t \leq 1/2$
to bound the rest.
So
\begin{align}
\log(1-t) &= 0 - t - t^2/2 -t^3/3 - ... \nonumber\\
&\geq -t -t^2/2(1 + 1/2 + 1/4 + ...) = -t - t^2 \nonumber
\end{align}
and we have
\begin{align}
\log \left(n^m\left(1-\frac{m}{n}\right)^r \right)
&= m \log n + r \log\left(1-\frac{m}{n}\right) \nonumber \\
& \geq m \log n - r m /n - r m/n^2. \nonumber
\end{align}

\todo{why?}
Our assumption that $ne^{r/n} \rightarrow \lambda$ implies that
\begin{align}
r = n \log n - n \log \lambda + o(n) \nonumber
\end{align}
so $r(m/n)^2 \rightarrow 0$.
Then
\begin{align}
rm/n &=m \log n - m \log \lambda + o(n)m/n \nonumber \\
m \log \lambda &\rightarrow m \log n - rm/n \nonumber
\end{align}
and
\begin{align}
\log \left(n^m\left(1-\frac{m}{n}\right)^r \right)
&\geq m \log n - rm/n \nonumber
\end{align}
so
\begin{align}
\lim_{n\rightarrow \infty} n^m\left(1-\frac{m}{n}\right)^r
&\geq \lambda^m. \nonumber
\end{align}

Now we piece together the pieces.
To begin,
\begin{align}
{ n \choose m } (1-\frac{m}{n})^r \rightarrow \lambda^m /m! \nonumber
\end{align}
so
\begin{align}
p_0(r,n) \rightarrow \sum_{k=0}^\infty \lambda^k /k! = e^{-\lambda}. \nonumber
\end{align}

The result holds
\begin{align}
p_0(r,n-m) \rightarrow e^{-\lambda} \nonumber
\end{align}
for fixed $m$ since $(n-m)e^{r/(n-m)} \rightarrow \lambda$.

Finally,
\begin{align}
p_m(r,n) &= {n \choose m} (1-\frac{m}{n})^r p_0(r, n-m) \nonumber \\
&\rightarrow \frac{\lambda^m}{ m!} e^{-\lambda} \nonumber
\end{align}
so we have that $N_n \rightarrow \textrm{Pois}(\lambda)$ as claimed.

Set $r = n \log n + nx$.
Then 
\begin{align}
ne^{-r/n} &= n e ^{- \log n - x} \nonumber \\
&= n e {- \log n} e^ {- x} = e^{-x}\nonumber
\end{align}

Consider the cover time variable $\cov$.
Since $\cov \leq m$ if and only if the first $m$ steps
of the random walk visit all $n$ vertices, then by \cref{thm:pois},
\begin{align}
\pr(\cov \leq r) &\rightarrow e^{-e^{-x}} \nonumber \\
\pr(\cov - n \log n\leq nx) &\rightarrow e^{-e^{-x}} \nonumber
\end{align}
